---
title: "R Stats Club"
subtitle: "From SPSS to R"
author: "Damien DUPRE"
date: "17/04/2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, scipen = 999)
set.seed(123)
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  error = FALSE,
  eval = FALSE)

library(tidyverse)
```

# Translating SPSS to R

* https://mikebyrnehfhci.wordpress.com/2012/10/19/translating-spss-to-r-simple-anova/
* www.soc.univ.kiev.ua/sites/default/files/course/materials/translation_syntax.pdf

Describes commandes for:
* Descriptives
* Basic ANOVA
* Welch Adjustment
* Contrasts
* Posthocs

---

# {SPSStoR} package

https://www.r-bloggers.com/spss-to-r-an-r-package-to-convert-spss-syntax/

```{r}
devtools::install_github("lebebr01/SPSStoR")
library(SPSStoR)
```

1. Save your spss syntax in a text file
2. Include `GET FILE='path/to/your/file.sav'.` at the beinging
3. The end of each command should finish with a dot.
4. Run the syntax with the function `spss_to_r(file = "path/to/your/syntax.txt")`

Example

```{r}
spss_to_r(file = "./RSC_5_from_spss_to_r/unianova_spss.txt")
```

---

But...

It doesn't work properly, it just gives an idea of what to do.

If you want to perform an anova, you have to tidy the output of the spss_to_r

```{r}
# mod_1 <- lm(LOG_TENURE ~ WAVE_DUMMY + SENIORITY_DUMMY + factor(MOTIVES) + factor(TSEX) + factor(PSEX) + factor(MOTIVES) + : + factor(TSEX) + factor(MOTIVES) + : + factor(PSEX) + factor(TSEX) + : + factor(PSEX) + factor(MOTIVES) + : + factor(TSEX) + : + factor(PSEX), data = x)

mod_1 <- lm(LOG_TENURE ~ WAVE_DUMMY + SENIORITY_DUMMY + MOTIVES * TSEX * PSEX, data = x)
car::Anova(mod_1, type = 3) 
```

---
class: center, middle

# Thank you for your attention
don't hesitate to send me an email: damien.dupre@dcu.ie
