---
title: "R Stats Club"
subtitle: "Mokken Scale Analysis"
author: "Damien DUPRE"
date: "08/05/2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, scipen = 999)

knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed,
  message = FALSE, 
  warning = FALSE, 
  error = FALSE,
  eval = FALSE)

library(tidyverse)
```

# Mokken Scale Analysis in R

Mokken scale analysis (MSA) is a scaling procedure for both dichotomous and polytomous
items. 

MSA is a scaling technique for ordinal data and mainly used for scaling test and questionnaire data.

It consists of an item selection algorithm to partition a set of items into
Mokken scales and several methods to check the assumptions of two nonparametric item
response theory models: the monotone homogeneity model and the double monotonicity
model.

This method aims to find robust and unidimensional scales in sets of items, in a search
procedure initialised by taking the highest associated pairs of items, and continuing by
subsequent gradual inclusion of well-fitting items until a scale has been formed that does
not improve further when other items are added.

```{r}
# install.packages("mokken")

library(mokken)
```

---

# Important Functions in mokken

* coefH(data): Scalability coefficients
* aisp(data): Scale partitioning
* check.iio(data), check.pmatrix(data), check.restscore(data), check.monotonicity(data): Assumption checks
---

# Example

MSA was performed on Adjective Checklist (Gough and Heilbrun 1980) data, acl, which are available in mokken. 

The data (Vorst 1992) contain the scores of 433 students from the University of Amsterdam on 218 items from a Dutch version of the Adjective Checklist. 

Each item is an adjective with five ordered answer categories (0 = completely disagree, 1 = disagree, 2 = neither agree nor disagree, 3 = agree, 4 = completely agree). 

Form each adjective, the respondents must consider to what degree it describes their personality, and mark the answer category that fits best to this description.

```{r}
data(acl)
```

---

# Scalability Coefficients

If H = 1, the you have a perfect scale.

Rules of thumb:
* If H < .3 the items are unscalable.
* If .3 < H < .4, a scale is weakly scalable.
* If .4 < H < .5, a scale is moderately scalable.
* If .5 < H, a scale is strongly scaleable.

```{r}
acl.com <- acl[,1:10]

acl.com

coefH(acl.com)
```

---

# Algorithm for the automated Item Selection Procedure or AISP

There are 2 that have been implemented in mokken using the default lower bound c = .3 results in two Mokken scales:
* Hierarchical clustering algorithm:
    - Starts by takeing 2 items having largest value of Hij
(significantly different from 0).
    - Items adding that meet criteria until no more can be added.
    - Takes unselected items and starts another Mokken scale
    - Continues until no more scales can be created.
* Genetic algorithm

```{r}
aisp(acl.com)
aisp(acl.com, search="ga")
```

---

# References

Mokken Scale Analysis in R - Journal of Statistical Software
https://www.jstatsoft.org/article/view/v020i11/v20i11.pdf

New Developments in Mokken Scale Analysis in R 
https://www.jstatsoft.org/index.php/jss/article/view/v048i05/v48i05.pdf

Homepage Andries van der Ark
https://sites.google.com/a/tilburguniversity.edu/avdrark/mokken

Item Response Theory in R
http://statmath.wu.ac.at/people/trusch/IMPS2017/Exercises/solutions/Exercise_05.html

---
class: center, middle

# Thank you for your attention
don't hesitate to send me an email: damien.dupre@dcu.ie
