---
title: "R Stats Club"
subtitle: "t-test, ANOVA and linear regression"
author: "Damien DUPRE"
date: "27/03/2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{css, echo=FALSE}
.remark-code {
  font-size: 12px;
  overflow-x: visible;
}
```


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, scipen = 999)
set.seed(123)
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  error = FALSE)
```

# Statistical tests


| test             | DV         | IV           | baseR   |
|------------------|------------|--------------|---------|
| t-test           |1 continuous|1 categorical | t.test()|
| ANOVA            |1 continuous|1+ categorical| aov()   |
| linear regression|1 continuous|1+ continuous | lm()    |

Note: t-test and anova are special cases of linear models so they can be performed using the `lm()` function as well (see following slides)

https://lindeloev.github.io/tests-as-linear/ for more informations

---

# Create a dataframe

```{r}
gender <- c(rep("male",50), rep("female",50))
# create an object gender which combines:
## the word male repeated 50 times 
## and the word female repeated 50 times

education <- c(rep("second",33), rep("undergrad",33), rep("master",34))
# create an object education which combines:
## the word second repeated 33 times, 
## the word undergrad repeated 33 times 
## and the word master repeated 34 times

age <- sample(15:50, 100, replace = TRUE)
# create an object age which select a number between 15 and 50, 100 times

math_results <- rnorm(n = 100, mean = 50, sd = 25)
# create an object math_results which creates an array of 

data_raw <- data.frame(gender, education, age, math_results)
# create a dataframe object with 4 columns: gender, education, age, math_results

str(data_raw) # check the structure of the dataframe
```

---
class: center, middle

# t-test with base R

---

# t-test

```{r}
res_ttest <- t.test(math_results ~ gender, data = data_raw)

res_ttest
```

---

# t-test with lm()

```{r}
res_lm_ttest <- lm(math_results ~ gender, data = data_raw)

summary(res_lm_ttest)
```

---
class: center, middle

# ANOVA with base R

---

# ANOVA

```{r}
res_anova <- aov(math_results ~ gender + education, data = data_raw)

summary(res_anova)
```

---

# ANOVA with lm()

```{r}
res_lm_anova <- lm(math_results ~ gender + education, data = data_raw)

# summary(res_lm_anova)

drop1(res_lm_anova, test = "F")
```

---
class: center, middle

# linear regression with base R

---

# linear regression

```{r}
res_lm <- lm(math_results ~ age, data = data_raw)

summary(res_lm)
```

---

# linear regression (main effects)

```{r}
res_lm <- lm(math_results ~ age + gender, data = data_raw)

# summary(res_lm)

drop1(res_lm, test = "F")
```

---

# linear regression (interaction effect)

```{r}
res_lm <- lm(math_results ~ age : gender, data = data_raw)

# summary(res_lm)

drop1(res_lm, test = "F")
```

---

# linear regression (main and interaction effects)

```{r}
res_lm <- lm(math_results ~ age * gender, data = data_raw)

summary(res_lm)
```

---
class: center, middle

# statistics with tadaatoolbox

```{r}
# install.packages("tadaatoolbox") # only once

library(tadaatoolbox)
```

---

# t-test with tadaatoolbox

```{r}
tadaa_t.test(data = data_raw, response = math_results, group = gender, print = "markdown")
```

---

# ANOVA with tadaatoolbox

```{r}
tadaa_aov(data = data_raw, math_results ~ gender + education, print = "markdown")
```

---
class: center, middle

# statistics with jmv

```{r}
# install.packages("jmv") # only once

library(jmv)
```

---

# t-test with jmv

```{r}
ttestIS(formula = math_results ~ gender, data = data_raw)
```

---

# ANOVA with jmv

```{r}
ANOVA(formula = math_results ~ gender + education, data = data_raw)
```

---
class: center, middle

# Thank you for your attention
don't hesitate to send me an email: damien.dupre@dcu.ie
